---
title       : "Data Visualization: ggloop vs tidyr & purrr"
date        : "12-12-2016"
author      : "Luke Smith"
description : "A few tricks to create multiple histograms."
tags        : [r, graphics, ggloop, tidyr, purrr]
---

A while back, I stumbled upon Simon Jackson's blog and some of his interesting methods [[1]](https://drsimonj.svbtle.com/quick-plot-of-all-variables) [[2]](https://drsimonj.svbtle.com/plot-some-variables-against-many-others) for creating a large number of histograms and point chars by using the `purrr`, `tidyr`, and `ggplot2` packages. These methods (and their results) are very useful, so I'd like to see if I can recreate them with `ggloop` and `ggplot2`.

## Method 1: Faceting points

#### tidyr

```{r purrr.tidyr1, message=FALSE}
library(tidyr)
library(ggplot2)

p1 <- mtcars %>%
  gather(-mpg, -hp, -cyl, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = mpg, color = hp, shape = factor(cyl))) +
    geom_point() +
    facet_wrap(~ var, scales = "free") +
    theme_bw()

print(p1)
```

#### ggloop

```{r ggloop1}
library(ggloop)

p2 <- mtcars %>%
  ggloop(aes_loop(disp:carb, mpg, color = hp, shape = factor(cyl))) %L+%
  geom_point() %L+%
  theme_bw()
```

Of course, there is a major problem with the `ggloop` approach - and that is that there is no way to view all plots at once, at least without having to use a secondary function. I am guilty of being too lazy to incorporate such a function into `ggloop`. However, there is a template available via Winston Chang's [Cookbook for R](http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/):


```{r multiplot}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

With that function, I can print the results of the `ggloop` function:

```{r ggloop1.print}
multiplot(plotlist = p2[[1]], cols =3)
```

#### Method 1: Conclusion
I should note that using `ggloop` with `multiplot()` is the slowest approach. I also prefer to see the x-value (`am`, `disp`, etc) at the top of each graph. The `ggloop` approach also produces individual graphs for each plot but for some reason cuts off the neam of the first legend in the top row of plots.

## Method 2:

## Method X:
How to overlay facets with `tidyr` method.