---
title  : "The Decline and Fall of the World Series Winner"
date   : "2017-02-25"
author : "Luke Smith"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::read_chunk("~/R/misc/baseball/ws_analysis.R")

# Plot set up
my_red        <- "#ce1141"
my_red2       <- "#a33c56"
my_lightred   <- "#fabdcd"
my_darkred    <- "#870b2b"
my_green      <- "#008348"
my_green2     <- "#1e6545"
my_lightgreen <- "#00be68"
my_darkgreen  <- "#00371e"
my_blue       <- "#4fa8ff"
my_blue2      <- "#1141ce"
my_orange     <- "#e15c39"
my_orange2    <- "#834800"
```

## It's not like it use to be
Take a look at the following chart which shows the regular season winning percentage of World Series winners from 1903 to 2015.

```{r Load_Dependencies, include=FALSE}
```
```{r Make_Data, include=FALSE}
```
```{r Plot_WSWinner, echo=FALSE, message=FALSE}
```
```{r __Plot_WSWinner, echo=FALSE, message=FALSE}
WS$Winner +
  geom_point(size = 1.75) +
  geom_line(size = 1) +
  geom_smooth(se = FALSE) +
  scale_y_continuous(minor_breaks = NULL) +
  scale_x_continuous(minor_breaks = NULL) +
  theme_minimal()
```


The decline in the regular-season winning percentage of World Series winner is also a trait of the loser, too.

```{r Plot_WSLoser, echo=FALSE, message=FALSE}
```
```{r __Plot_WSLoser, echo=FALSE, message=FALSE}
WS$Loser
```

```{r inline_calc1, include=FALSE}
filt <- . %>%
  group_by_(WSTeams2, cond = .) %>%
  summarize(n = n()) %>%
  filter(cond == TRUE) %>%
  select(n) %>%
  .[[1]]

W    <- "Y > N"  %>% filt()
L    <- "Y < N"  %>% filt()
G    <- "Y != N" %>% filt()
ties <- "Y == N" %>% filt()
```

The team with the better winning percentage has only a `r W`-`r L` record out of the `r G` World Series where one team had a greater winning percentage. In `r ties` games, both teams had identical winning percentages.

In the color-coded chart below, you can see both the difference in winning percentages and where the eventual World Series winner (blue) and loser (red) stood.

```{r Plot_WS_WL_Diff1, echo=FALSE, message=FALSE}
```
```{r __Plot_WS_WL_Diff1, echo=FALSE, message=FALSE}
WS$Teams #+
  # scale_y_continuous(minor_breaks = NULL) +
  # scale_x_continuous(minor_breaks = NULL) +
  # theme_minimal()
```
## Conclusion - if any

There is probably a reason to all of the this (i.e. game rule changes, league expansion, changes in player contracts, etc). However, that will have to be left for another post.

<br>
<br>