---
title : "The tidyverse vs base R: Example 2"
author: "Luke Smith"
date  : "03-02-2017"
description: "Bring order to a chaotic universe."
tags: [r, tidyverse, base R, data wrangling]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# library(RODBC)
# db <- odbcDriverConnect('driver=SQL Server;server=DAGOBAH\\SQLEXPRESS1;database=lahman;trusted_connection=true')
# knitr::opts_chunk$set(connection = "db")
```

Besides being occasionally better than base R, the tidyverse can also overcome some of the verbosity and complexity of SQL.

## dplyr vs base R vs SQL
### Finding the Youngest Player-Debut For Every Season
For example, let's find the youngest player to have debut each year in Major League Baseball (MLB) using the Lahman database (which the R version can be found [here][R_Lahman] and the SQL version [here][SQL_Lahman]).

[R_Lahman]: https://cran.r-project.org/package=Lahman
[SQL_Lahman]: http://www.seanlahman.com/baseball-archive/statistics/

```{r data_Lahman, message=FALSE}
library(Lahman)

Master <- tibble::as_tibble(Master)

# The Lahman master table
print(Master)
```

<br>

All three methods will be presented below.

#### tidyverse

```{r tidyverse_Lahman, message=FALSE}
library(tidyverse)
library(lubridate) # Included in the install; must be loaded separately

Master2 <- Master %>%
  filter(!is.na(birthDate) & !is.na(debut)) %>%
  mutate(debut      = as.Date(debut),
         debut_year = year(debut),
         debut_age  = debut - birthDate)

tidyverse_result <- Master2 %>%
  group_by(debut_year = year(debut)) %>%
  filter(debut_age == min(debut_age)) %>%
  arrange(debut_year) %>%
  ungroup()
```

<br>

#### base R
And now the base R solution which uses assignment (`<-`) in place of `dplyr::mutate()`; `split()` in place of `dplyr::group_by()`; and single-bracket (`[`) subsetting in place of `dplyr::filter()`.

```{r base_R_Lahman, message=FALSE}
keep_this <- !is.na(Master$birthDate)
keep_that <- !is.na(Master$debut)

Master2 <- Master[keep_this & keep_that, ]
debut_split <- strsplit(x     = as.character.Date(Master2$debut),
                            split = "-",
                            fixed = TRUE)
Master2$debut_year <- vapply(debut_split, `[`, character(1), 1)
Master2$debut_age  <- as.Date(Master2$debut) - Master2$birthDate

by_debut_year <- split(Master2, Master2$debut_year)

get_min <- function(x) x[x$debut_age == min(x$debut_age), ]

ydbR <- lapply(by_debut_year, get_min)
ydbR <- Reduce(function(...) merge(..., all = TRUE), ydbR)
base_R_result <- ydbR[order(ydbR$debut_year), ]
```

<br>

And last, the SQL solution, using T-SQL within SQL Server. I've decided to hide the actual function used to retrieve the data in favor of showing you the syntax-highlighted SQL statement.

```{r SQL_Lahman_connect, message=FALSE}
library(RODBC)

db <- odbcDriverConnect('driver=SQL Server;server=DAGOBAH\\SQLEXPRESS1;database=lahman;trusted_connection=true')
```
```{r SQL_Lahman, include=FALSE}
sql_result <- sqlQuery(db, "SELECT playerID, nameFirst, nameLast, debutAge, debut
FROM Master
INNER JOIN (
	SELECT MIN(debutAge) as debutAge, m.debutYear
	FROM (
		SELECT *,
			YEAR(debut) as debutYear,
			DATEDIFF(day, CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date), debut) as debutAge,
			CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date) as birthDate
		FROM Master
		WHERE birthDay IS NOT NULL AND
			birthMonth IS NOT NULL AND
			birthYear IS NOT NULL AND
			debut IS NOT NULL
		) as m
	GROUP BY m.debutYear
	) as n
ON DATEDIFF(day, CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date), debut) = n.debutAge AND
	YEAR(debut) = n.debutYear
ORDER BY YEAR(debut)
;")
```

```sql
SELECT playerID, nameFirst, nameLast, debutAge, debut
FROM Master
INNER JOIN (
	SELECT MIN(debutAge) as debutAge, m.debutYear
	FROM (
		SELECT *,
			YEAR(debut) as debutYear,
			DATEDIFF(day, CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date), debut) as debutAge,
			CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date) as birthDate
		FROM Master
		WHERE birthDay IS NOT NULL AND
			birthMonth IS NOT NULL AND
			birthYear IS NOT NULL AND
			debut IS NOT NULL
		) as m
	GROUP BY m.debutYear
	) as n
ON DATEDIFF(day, CAST(CAST(birthYear as nchar) + '/' + CAST(birthMonth as nchar) + '/' + CAST(birthDay as nchar) as date), debut) = n.debutAge AND
	YEAR(debut) = n.debutYear
ORDER BY YEAR(debut)
;
```
<br>

Wow. I don't think you have to be much of a programmer, coder, hacker, or whatever to tell which one of the three you would rather __NOT USE__ on a daily basis.

```{r}
# Compare tidyverse to base R
anti_join(tidyverse_result, base_R_result, by = "playerID")

# Compare base R to tidyverse
anti_join(base_R_result, tidyverse_result, by = "playerID")

# Compare tidyverse to SQL
anti_join(tidyverse_result, sql_result, by = "playerID")

# Compare SQL to tidyverse
anti_join(tidyverse_result, sql_result, by = "playerID")
```

<br>
<br>